---
title: "CFU"
author: "Amy Fox"
output: html_document
---

This R-markdown plots the Colony Forming Unit (CFU) data to determine if the differences in bacterial burden between the Control (PBS) and Vaccinated (BCG) are statistically significant at the different time points.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the necessary packages
```{r warning= FALSE, message = FALSE}
library(readxl)
library(stats)
library(ggpubr)
library(broom)
library(purrr)
library(tidyverse)
```

Read in excel file with CFU data and clean.
Note: For the publication we are only analyzing lung data, so filter to lung.
Additional Note: The excel file contains the data organized with the following columns: Group (experimental group), Number (the mouse number within a group), Organ (organ from which the data is collected), Timepoint (day in which the data was taken post-infection), CFU (log-10 colony forming units).
```{r}

lung_CFU <- read_xlsx("../Data/M.smeg_CFU.xlsx") %>%
  rename(Day = Timepoint) %>%
  group_by(Organ, Day) %>%
  mutate(Group = replace(Group, Group == "Group1", "PBS")) %>%
  mutate(Group = replace(Group, Group == "Group2", "BCG")) %>%
  mutate(Group = replace(Group, Group == "Group3", "smegmatis")) %>%
  mutate(Group = replace(Group, Group == "Group4", "BCG + smegmatis")) %>% 
  mutate(Group = factor(Group, levels = c("PBS", "BCG", "smegmatis", "BCG + smegmatis"))) %>%
  dplyr::filter(Group == "PBS" | Group == "BCG") %>%
  filter(Organ == "Lung")

```

Plot the lung CFU data and run t tests
```{r fig.width = 3, fig.height = 1.2}

viridis_colors <- c("#56C667FF", "#3F4788FF")

lung_CFU <- lung_CFU %>%
  mutate(Group = str_replace(Group, "PBS", "Control"),
         Group = str_replace(Group, "BCG", "Vaccinated"))

ggplot(lung_CFU, aes(x = Group, y = CFU, color = Group)) +
  geom_jitter(width = 0.2) +
  facet_wrap("Day", ncol = length(unique(lung_CFU$Day)), 
             labeller = label_both) +
  scale_color_manual(values = viridis_colors) +
  stat_compare_means(method = "t.test", label.y = 7.5) +
  xlab("Mouse Group") +
  ylab("log 10 CFU") +
  coord_cartesian(ylim = c(2,8)) +
  ggtitle("Lung Bacterial Load") +
  theme_bw()

```

```{r}
sessionInfo()
```