---
title: "CFU"
author: "Amy Fox"
output: html_document
---

This R-markdown plots the Colony Forming Unit (CFU) data to determine if the differences in bacterial burden between the Control (PBS) and Vaccinated (BCG) are statistically significant at the different time points.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the necessary packages
```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(stats)
library(ggpubr)
library(broom)
library(purrr)
library(tidyr)
```

Read in excel file with CFU data and clean.
Note: For the publication we are only analyzing lung data, so filter to lung.
```{r}

lung_CFU <- read_xlsx("../Data/M.smeg_CFU.xlsx") %>%
  rename(Day = Timepoint) %>%
  group_by(Organ, Day) %>%
  mutate(Group = replace(Group, Group == "Group1", "PBS")) %>%
  mutate(Group = replace(Group, Group == "Group2", "BCG")) %>%
  mutate(Group = replace(Group, Group == "Group3", "smegmatis")) %>%
  mutate(Group = replace(Group, Group == "Group4", "BCG + smegmatis")) %>% 
  mutate(Group = factor(Group, levels = c("PBS", "BCG", "smegmatis", "BCG + smegmatis"))) %>%
  dplyr::filter(Group == "PBS" | Group == "BCG") %>%
  filter(Organ == "Lung")

```

Plot the lung CFU data and run t tests
```{r fig.width = 3, fig.height = 1.2}

viridis_colors <- c("#56C667FF", "#3F4788FF")

lung_CFU <- lung_CFU %>%
  mutate(Group = str_replace(Group, "PBS", "Control"),
         Group = str_replace(Group, "BCG", "Vaccinated"))

ggplot(lung_CFU, aes(x = Group, y = CFU, color = Group)) +
  geom_jitter(width = 0.2) +
  facet_wrap("Day", ncol = length(unique(lung_CFU$Day)), 
             labeller = label_both) +
  scale_color_manual(values = viridis_colors) +
  stat_compare_means(method = "t.test", label.y = 7.5) +
  xlab("Mouse Group") +
  ylab("log 10 CFU") +
  coord_cartesian(ylim = c(2,8)) +
  ggtitle("Lung Bacterial Load") +
  theme_bw()

```

