---
title: "Human_clinical_remove_sc"
author: "Amy Fox"
output: html_document
---

This dataset includes human whole blood from a healthy individual collected on an LSR Fortessa II flow cytometer.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, the packages must be loaded.
```{r warning= FALSE, message = FALSE}
library(openCyto)
library(flowClust)
library(data.table)
library(flowCore)
library(ggcyto)
library(FlowSOM)
library(pheatmap)
library(stringr)
library(readxl)
library(gridExtra)
library(ggcorrplot)
library(tidyverse)
library(scales)
library(viridis)
library(kableExtra)
library(ggpubr)
```

I have created a few functions to make it easier to analyze the data.

- clean_sc_after_gate:renames the columns; it also removes SSC boundary effects
- .findpeaks: copied from <https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R> to use in later mindensity function
- .findvalleys: copied from <https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R> to use in later mindensity function
- gate_mindensity_amy: this is a function that I adapted from <https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R> the difference between the original mindensity function and mine is that the original only works with a flowframe, while mine performs the same function on a dataframe. Mine also prints out the cutpoint so we can see it.
- statSmoothFunc: copied from <https://gist.github.com/kdauria/524eade46135f6348140> to add r^2 to correlation plots
- stat_smooth_func: copied from <https://gist.github.com/kdauria/524eade46135f6348140> to add r^2 to correlation plots
- as.numeric.factor: convert the factors for the marker names into numeric for plotting
```{r}

clean_sc_after_gate <- function(fcs_file) {
  fcs_file %>%
  select(ends_with(".A"), -`FSC.A`, `SSC.A`, filename) %>%
  dplyr::rename(`HLA_DR` = "FITC.A",
         `CD19` = "PE.A",
        `CD3` =  "PE.Cy5.A",       
         `CD45` = "PE.Cy7.A",
         `CD27` = "APC.A")  %>%
   na.omit()%>%
  dplyr::filter(`SSC.A` != max(`SSC.A`))

}

.between_interval <- function(x, interval) {
  x <- x[findInterval(x, interval) == 1]
  if (length(x) == 0) {
    x <- NA
  }
  x
}

# this function is from OpenCyto Github RGLab https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R
.find_peaks <- function(x, y = NULL, num_peaks = NULL, adjust = 2, plot = FALSE, ...) {
  x <- as.vector(x)

  if (length(x) < 2) {
    warning("At least 2 observations must be given in 'x' to find peaks.")
    return(NA)
  }

  if (is.null(y)) {
    dens <- density(x, adjust = adjust, ...)
  } else {
    y <- as.vector(y)
    if (length(x) != length(y)) {
      stop("The lengths of 'x' and 'y' must be equal.")
    }
    dens <- list(x = x, y = y)
  }

  # Discrete analogue to a second derivative applied to the KDE. See details.
  second_deriv <- diff(sign(diff(dens$y)))
  which_maxima <- which(second_deriv == -2) + 1

  # The 'density' function can consider observations outside the observed range.
  # In rare cases, this can actually yield peaks outside this range.  We remove
  # any such peaks.
  which_maxima <- which_maxima[findInterval(dens$x[which_maxima], range(x)) == 1]

  # Next, we sort the peaks in descending order based on the density heights.
  which_maxima <- which_maxima[order(dens$y[which_maxima], decreasing = TRUE)]
  
  # Returns the local maxima. If there are none, we return 'NA' instead.
  if (length(which_maxima) > 0) {
    peaks <- dens$x[which_maxima]
    if (is.null(num_peaks) || num_peaks > length(peaks)) {
      num_peaks <- length(peaks)
    }
    peaks <- peaks[seq_len(num_peaks)]
  } else {
    peaks <- NA
  }
  
  peaks <- data.frame(x = peaks, y = dens$y[which_maxima][seq_len(num_peaks)])
  if(plot){
    plot(dens, main = paste("adjust =" ,  adjust))
    points(peaks, ,col = "red")  
  }
  
  peaks  
}

# this function is from OpenCyto Github RGLab https://github.com/RGLab/openCyto/blob/0ec1218427627b9ee092f8d39e6ab46d46e7e2f8/R/bayes-flowClust.R
.find_valleys <- function(x, y = NULL, num_valleys = NULL, adjust = 2, ...) {

  x <- as.vector(x)

  if (length(x) < 2) {
    warning("At least 2 observations must be given in 'x' to find valleys.")
    return(NA)
  }
  
  if (is.null(y)) {
    dens <- density(x, adjust = adjust, ...)
  } else {
    y <- as.vector(y)
    if (length(x) != length(y)) {
      stop("The lengths of 'x' and 'y' must be equal.")
    }
    dens <- list(x = x, y = y)
  }

  # Discrete analogue to a second derivative applied to the KDE. See details.
  second_deriv <- diff(sign(diff(dens$y)))
  which_minima <- which(second_deriv == 2) + 1

  # The 'density' function can consider observations outside the observed range.
  # In rare cases, this can actually yield valleys outside this range. We remove
  # any such valleys.
  which_minima <- which_minima[findInterval(dens$x[which_minima], range(x)) == 1]

  # Next, we sort the valleys in descending order based on the density heights.
  which_minima <- which_minima[order(dens$y[which_minima], decreasing = FALSE)]

  # Returns the local minima. If there are none, we return 'NA' instead.
  if (length(which_minima) > 0) {
    valleys <- dens$x[which_minima]
    if (is.null(num_valleys) || num_valleys > length(valleys)) {
      num_valleys <- length(valleys)
    }
    valleys <- valleys[seq_len(num_valleys)]
  } else {
    valleys <- NA
  }
  valleys
}


gate_mindensity_amy <- function(df, channel, filterId = "", positive = TRUE,
                       pivot = FALSE, gate_range = NULL, min = NULL, max = NULL,
                       peaks = NULL, ...) {
  
  if (missing(channel) || length(channel) != 1) {
    stop("A single channel must be specified.")
  }

  # Filter out values less than the minimum and above the maximum, if they are
  # given.

  # I altered this to take in a dataframe rather than fcs file
  x <- df[, channel]
  
  if(is.null(peaks))
    peaks <- .find_peaks(x)[, "x"]
  
  if (is.null(gate_range)) {
    gate_range <- c(min(x), max(x))
  } else {
    gate_range <- sort(gate_range)
  }

  
  
  # In the special case that there is only one peak, we are conservative and set
  # the cutpoint as min(x) if 'positive' is TRUE, and max(x) otherwise.
  if (length(peaks) == 1) {
    cutpoint <- ifelse(positive, gate_range[1], gate_range[2])
  } else {
    # The cutpoint is the deepest valley between the two peaks selected. In the
    # case that there are no valleys (i.e., if 'x_between' has an insufficient
    # number of observations), we are conservative and set the cutpoint as the
    # minimum value if 'positive' is TRUE, and the maximum value otherwise.
    valleys <- try(.find_valleys(x), silent = TRUE)
    valleys <- .between_interval(x = valleys, interval = gate_range)

    if (any(is.na(valleys))) {
    #FIXME:currently it is still returning the first peak,
    #we want to pass density instead of x_between to 'min'
    #because x_between is the signal values
      cutpoint <- ifelse(positive, gate_range[1], gate_range[2])
    } else if (length(valleys) == 1) {
      cutpoint <- as.vector(valleys)
    } else if (length(valleys) > 1) {
      # If there are multiple valleys, we determine the deepest valley between
      # the two largest peaks.
      peaks <- sort(peaks[1:2])
      cutpoint <- .between_interval(valleys, peaks)[1]

      # If none of the valleys detected are between the two largest peaks, we
      # select the deepest valley.
      if (is.na(cutpoint)) {
        cutpoint <- valleys[1]
      }      
    }
  }
  gate_coordinates <- ifelse(positive, list(c(cutpoint, Inf)), list(c(-Inf, cutpoint)))
  
  names(gate_coordinates) <- channel
  
  rectangleGate(gate_coordinates, filterId = filterId)
  
  # I added in the print(cutpoint)
  print(cutpoint)
  
}

# taken from https://gist.github.com/kdauria/524eade46135f6348140
stat_smooth_func <- function(mapping = NULL, data = NULL,
                        geom = "smooth", position = "identity",
                        ...,
                        method = "auto",
                        formula = y ~ x,
                        se = TRUE,
                        n = 80,
                        span = 0.75,
                        fullrange = FALSE,
                        level = 0.95,
                        method.args = list(),
                        na.rm = FALSE,
                        show.legend = NA,
                        inherit.aes = TRUE,
                        xpos = NULL,
                        ypos = NULL) {
  layer(
    data = data,
    mapping = mapping,
    stat = StatSmoothFunc,
    geom = geom,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      method = method,
      formula = formula,
      se = se,
      n = n,
      fullrange = fullrange,
      level = level,
      na.rm = na.rm,
      method.args = method.args,
      span = span,
      xpos = xpos,
      ypos = ypos,
      ...
    )
  )
}

StatSmoothFunc <- ggproto("StatSmooth", Stat,
                      
                      setup_params = function(data, params) {
                        # Figure out what type of smoothing to do: loess for small datasets,
                        # gam with a cubic regression basis for large data
                        # This is based on the size of the _largest_ group.
                        if (identical(params$method, "auto")) {
                          max_group <- max(table(data$group))
                          
                          if (max_group < 1000) {
                            params$method <- "loess"
                          } else {
                            params$method <- "gam"
                            params$formula <- y ~ s(x, bs = "cs")
                          }
                        }
                        if (identical(params$method, "gam")) {
                          params$method <- mgcv::gam
                        }
                        
                        params
                      },
                      
                      compute_group = function(data, scales, method = "auto", formula = y~x,
                                               se = TRUE, n = 80, span = 0.75, fullrange = FALSE,
                                               xseq = NULL, level = 0.95, method.args = list(),
                                               na.rm = FALSE, xpos=NULL, ypos=NULL) {
                        if (length(unique(data$x)) < 2) {
                          # Not enough data to perform fit
                          return(data.frame())
                        }
                        
                        if (is.null(data$weight)) data$weight <- 1
                        
                        if (is.null(xseq)) {
                          if (is.integer(data$x)) {
                            if (fullrange) {
                              xseq <- scales$x$dimension()
                            } else {
                              xseq <- sort(unique(data$x))
                            }
                          } else {
                            if (fullrange) {
                              range <- scales$x$dimension()
                            } else {
                              range <- range(data$x, na.rm = TRUE)
                            }
                            xseq <- seq(range[1], range[2], length.out = n)
                          }
                        }
                        # Special case span because it's the most commonly used model argument
                        if (identical(method, "loess")) {
                          method.args$span <- span
                        }
                        
                        if (is.character(method)) method <- match.fun(method)
                        
                        base.args <- list(quote(formula), data = quote(data), weights = quote(weight))
                        model <- do.call(method, c(base.args, method.args))
                        
                        m = model
                        eq <- substitute(~~italic(r)^2~"="~r2, 
                                         list(a = format(coef(m)[1], digits = 3), 
                                              b = format(coef(m)[2], digits = 3), 
                                              r2 = format(summary(m)$r.squared, digits = 3)))
                        func_string = as.character(as.expression(eq))
                        
                        if(is.null(xpos)) xpos = min(data$x)*0.9
                        if(is.null(ypos)) ypos = max(data$y)*0.9
                        data.frame(x=xpos, y=ypos, label=func_string)
                        
                      },
                      
                      required_aes = c("x", "y")
)

viridis_colors <- c("#3F4788FF", "#56C667FF")

# need to convert the factors in the extracted_pheno to numeric for plotting
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
```

## Data cleaning
Initial gating of sample data
```{r fig.width = 9, fig.height=5, warning=FALSE}
# read in the gating strategy
ws <- list.files("../data/flow/Human_Clinical", pattern = "human_clinical_gating.csv", full = TRUE)

# only for viewing that read in 
dtTemplate <- fread(ws) 

initial_gate <- gatingTemplate(ws) 

fcsFile <- list.files("../data/flow/Human_Clinical/", 
                       pattern = ".fcs", full = TRUE)

# ncdfFlowset object contains row names with the individual samples and column names with the markers/parameters used in the flow cytometer
ncfs <- read.ncdfFlowSet(fcsFile) 

# apply gating set
gs <- GatingSet(ncfs)

# gate the samples
gt_gating(initial_gate, gs)

# View the gates
plotGate(gs[[1]])

# Pull out the data
gated_data <- gs_pop_get_data(gs, "leukocytes") %>% 
  as.flowSet() 

# initialize the gated data df
all_gated_df = data.frame()

# logicle transform of data in the listed channels 
# convert each of the FMO files to a data frame, adding on the filename
# bind all of the FMO files into one dataframe (though there is only 1 file)
for (i in 1:length(gated_data)) {
  lgcl <- estimateLogicle(gated_data[[i]], channels = c( "FITC-A", "PE-A",
                                                 "PE-Cy5-A", "PE-Cy7-A", "APC-A"))
  
  gated_data_test <- transform(gated_data[[i]], lgcl) %>%
    exprs() %>%
    data.frame() %>%
    mutate(filename = sampleNames(gated_data[i])) 

  all_gated_df <- rbind(all_gated_df, gated_data_test)
}


all_gated_for_plot <- all_gated_df %>%
  clean_sc_after_gate()


 

```

## Clear separation of populations --> Gate with minimum density between + and - populations
```{r fig.width = 5.25, fig.height = 3}

separation_ggplot <- function(df, x_axis, x_title) {
  ggplot(df, aes(x = x_axis, y = SSC.A)) +
   geom_hex(bins = 100) +
   scale_fill_viridis_c() +
   xlab(x_title) +
   ylab("SSC-A") +
   geom_vline(xintercept = gate_mindensity_amy(df, x_title), size = .7) +
  theme_gray()+
  theme(axis.text = element_text(size =12),
        axis.title = element_text(size = 20),
        legend.title=element_text(size=12),
        legend.text=element_text(size=rel(.75)),
        legend.key.size = unit(.5, "cm"))
}

p1 <- separation_ggplot(all_gated_for_plot, all_gated_for_plot$HLA_DR, "HLA_DR")
p2 <- separation_ggplot(all_gated_for_plot, all_gated_for_plot$CD19, "CD19")
p3 <- separation_ggplot(all_gated_for_plot, all_gated_for_plot$CD3, "CD3")
p4 <- separation_ggplot(all_gated_for_plot, all_gated_for_plot$CD45, "CD45")
p5 <- separation_ggplot(all_gated_for_plot, all_gated_for_plot$CD27, "CD27")

fig_6a <- grid.arrange(p1, p2, p3, p4, p5, 
             nrow = 2)

fig_6a 

ggsave("../figures/fig_6a.jpeg", fig_6a, dpi = 300)
```

## Feature engineer the data
Feature cut the data to get all of the possible populations for each file with the cell count and number of cells. Feature engineering is based on the minimum density between the single color peaks.

```{r}

all_gated <- all_gated_df %>%
  mutate(filename= str_replace_all(filename, "\\s", "_")) %>%
  select(ends_with(".A"), -`FSC.A`, filename) %>%
  dplyr::rename(`HLA_DR` = "FITC.A",
         `CD19` = "PE.A",
        `CD3` =  "PE.Cy5.A",       
         `CD45` = "PE.Cy7.A",
         `CD27` = "APC.A")  %>%
  na.omit() %>%
  dplyr::filter(`SSC.A` != max(`SSC.A`)) %>%
  mutate(HLA_DR = cut(HLA_DR, breaks = c(min(HLA_DR), gate_mindensity_amy(all_gated_for_plot, "HLA_DR"), max(HLA_DR)), 
                       labels = c("0", "1")),
         CD19 = cut(CD19, breaks = c(min(CD19), gate_mindensity_amy(all_gated_for_plot, "CD19"), max(CD19)),
                       labels = c(0, 1)),
         CD3 = cut(CD3, breaks = c(min(CD3), gate_mindensity_amy(all_gated_for_plot, "CD3"), max(CD3)),
                       labels = c(0, 1)),
         CD45 = cut(CD45, breaks = c(min(CD45), gate_mindensity_amy(all_gated_for_plot, "CD45"),  max(CD45)), 
                    labels = c(0, 1)),
         CD27 = cut(CD27, breaks = c(min(CD27), gate_mindensity_amy(all_gated_for_plot, "CD27"), max(CD27)), 
                    labels = c(0, 1))) %>%
  mutate_at(vars(-group_cols()), list(~tidyr::replace_na(.,0))) %>%
  select(-`SSC.A`) %>%
  group_by(HLA_DR, CD19, CD3, CD45, CD27, filename) %>%
  dplyr::mutate(cell_no = n()) %>%
  unique() %>%
  ungroup() %>%
  dplyr::group_by(filename) %>%
  dplyr::mutate(total_count_by_file = sum(cell_no),
         percentage = (100*cell_no / total_count_by_file)) %>%
  mutate_if(is.factor, ~as.numeric.factor(.)) %>%
  ungroup() 

```

## Initial identification of populations plot
Data visualization

We first want to view all of the different cell phenotypes within the data, and then we can filter the data to see the ones that we're interested in, for example, CD45+ leukocytes.

```{r}
# to view all of the possible combinations
total_phenotypes <- all_gated %>%
  select(-filename, -percentage, -cell_no, -total_count_by_file) %>%
  group_by(HLA_DR, CD19, CD3, CD45, CD27) %>%
  unique() %>%
  ungroup() %>%
 select(CD45, CD3, CD19, HLA_DR, CD27)
 # mutate_all(~as.numeric.factor(.))

  
fig_6b <- pheatmap(total_phenotypes, cluster_rows = FALSE, cluster_cols = FALSE, 
         color = viridis_colors, show_rownames = FALSE, angle_col = 45, legend = F, fontsize_col = 16)

fig_6b

ggsave("../figures/fig_6b.jpeg", fig_6b, dpi = 300)

# gives the total number of populations
nrow(total_phenotypes) 

# view the specific cell phenotypes and percentages we're interested in
sample_populations <- all_gated %>%
  dplyr::filter(CD45 == 1) %>%
  select(-filename, -percentage, -cell_no, -total_count_by_file) %>%
  group_by(CD45, CD3, CD19, HLA_DR, CD27) %>%
  unique() %>%
  ungroup() %>%
  mutate(population = paste0("Pop", 1:length(CD45)))

```

## Filter to smaller subset of populations and plot
Data visualization

- phenotype plot

Want to add annotations to the phenotype plot based on expression of different markers. The phenotype plot is then arranged and split by the lineage (I should try to find a way to do this automatically).

```{r}
# Split populations based on T or B cells
my_colors <- list(lineage = c(`Other Leukocyte` = "yellow",  `B cell` = "#F8766D", 
                              `T cell` = "darkorchid1",`CD3+ B cells` = "#00B0F6"))


test_rearrange <- sample_populations %>%
  select(CD45, CD3, CD19, HLA_DR, CD27) %>%
  arrange(CD3, CD19) %>%
  mutate(population = paste0("Pop", c(1:nrow(sample_populations)))) %>%
  column_to_rownames("population") 
  #mutate_all(~as.numeric.factor(.)) %>%
  
annotation <- test_rearrange %>%
  mutate(lineage = ifelse(CD3 == 1 & CD19 == 0 , "T cell",
                          ifelse(CD3 == 0 & CD19 == 1, "B cell", 
                                 ifelse(CD3 == 1 & CD19 == 1, "CD3+ B cells", "Other Leukocyte")))) %>%
  select(lineage) %>%
  mutate(population = paste0("Pop", c(1:nrow(sample_populations)))) %>%
  mutate(lineage = as.factor(lineage)) %>%
  column_to_rownames("population")
  
fig_6c <- pheatmap(test_rearrange, cluster_rows = FALSE, cluster_cols = FALSE,
           labels_row = rownames(annotation), 
           annotation_row = annotation,  annotation_colors = my_colors,  color = viridis_colors,
           cellwidth = 15, cellheight = 15, angle_col = 45, legend = F,
           main = "Leukocytes")

fig_6c

ggsave("../figures/fig_6c.jpeg", fig_6c, dpi = 300)
```

Note I cannot do a correlation plot because I only have 1 sample (not multiple datapoints to see how similar populations are to each other)

```{r}
# convert the population rowname back to column so that we can use it for further analysis
test_rearrange <- test_rearrange %>%
  rownames_to_column("population")

# add back in the percentage data 
sample_populations_all_groups <- left_join(test_rearrange, all_gated) %>%
  select(population, percentage, -filename) %>%
  mutate_all(list(~replace_na(., 0)))
```

# Bar plot
Data visualization

- percentage of cells in each population

```{r}

fig_6d <- ggplot(sample_populations_all_groups, aes(x = factor(population, 
                                                     levels = c(paste0("Pop",
                                                                       1:nrow(sample_populations_all_groups)))), 
                                          y = percentage, fill = "pink")) +
  geom_bar(stat = "identity") +
  xlab("Populations") +
  ylab("Percent of Cells") +
  ggtitle("Cell Percentage of Total Live Leukocytes") +
  scale_color_manual(values = viridis_colors) +
  theme_gray() +
  theme(axis.text.x = element_text(angle = 45, size = 15, hjust = 1),
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 10),
        axis.title = element_text(size = 20),
        title = element_text(size = 20),
        legend.position = "none")

fig_6d

ggsave("../figures/fig_6d.jpeg", fig_6d, dpi = 300)
```

Confirm pipeline populations with manual gating
```{r}

# read in and clean manual gating percentage data
manual_gating <- read_xlsx("../data/flow/Human_Clinical/Manual_gating.xlsx") %>%
  dplyr::rename(`population` = "...1",
                percent_manual = "Manual percent") %>%
  na.omit()

# prepare the flow CFU data for joining with manual gating
sample_populations_all_groups <- sample_populations_all_groups %>%
  dplyr::rename(percent_pipeline = percentage)

# should add absolute value of the percent_pipeline-percent_manual for error (abs)
# join the manual and pipeline percentage data
comparison <- full_join(manual_gating, sample_populations_all_groups, by = c("population")) %>%
  mutate(difference = round(percent_pipeline - percent_manual, digits = 2)) %>%
  mutate(error = round(100*(percent_pipeline - percent_manual)/percent_manual, digits = 2)) %>%
  mutate(percent_manual = round(percent_manual, digits = 2),
         percent_pipeline = round(percent_pipeline, digits = 2)) %>%
  mutate(absolute_difference = abs(difference),
         average_difference = round(mean(absolute_difference), digits = 2)) %>%
  select(population, percent_manual, percent_pipeline, absolute_difference, average_difference) 

# print out the comparison data in a pretty table
fig_6e <- comparison[1:14,] %>%
  mutate(Population = population,
         `% Based On Manual Gating` = percent_manual,
         `% Based on Automated Pipeline` = percent_pipeline,
         `Absolute Difference` = absolute_difference,
         `Average Difference` = average_difference) %>%
  select(Population, `% Based On Manual Gating`, `% Based on Automated Pipeline`, `Absolute Difference`, `Average Difference`) %>%
  kable(align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 25) %>%
  row_spec(0:nrow(comparison), color = "black") 

fig_6e

save_kable(fig_6e, file = "../figures/fig_6e.jpeg")

# compare the two gating types with manual v. pipeline plot
fig_6f <- ggplot(comparison, aes(x = percent_pipeline, y = percent_manual)) +
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0, size = 1) +
  stat_cor(aes(x = percent_pipeline, y = percent_manual), data = comparison, "spearman", size = 7) +
  ggtitle("Human Clinical Data: Manual vs. Pipeline Percentage") +
  xlab("Pipeline Percent") +
  ylab("Manual Percent") +
  theme_classic() +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 15),
        title = element_text(size = 15)) 

fig_6f

ggsave("../figures/fig_6f.jpeg", fig_6f, dpi = 300)
```

```{r}
BiocManager::valid()
```